\section{Conclusion}

Supply Chain Security is an increasingly important factor of modern Software Development. Ensuring that companies adopt good practices in regards to the code they outsource is a pivotal step in ensuring the security requirements of software products. SBOMs \cite{article:concept:sbom-2} are an example of such a technology that promotes good secure software engineering practices.

\noindent In this paper, we analyzed 5 examples of the state-of-the-art with regards to SBOMs and conducted a critical review of these papers according to our own metrics.

\noindent According to our review, we believe that, although arguably the article with the least quality out of the 5 reviewed based on our metrics and subjective evaluation, the work done by Zahan et al. in \cite{article:sbom-required} is a good starting point towards higher SBOM expertise. Stalnaker et al. provide in \cite{article:software-bom} the greatest level of detail regarding common challenges practitioners and organizations might face when attempting to adopt SBOM best practices. When it comes to employing solutions to mitigate those challenges, the work done by Bi et al. in \cite{article:sboms-issues-solutions} can be a good reference point for the steps needed to solve any problems encountered.

\noindent Future work can expand on the methodology and results of this paper by analyzing more articles to provide a broader and more comprehensive look on the state of SBOM-related literature and by coming up with more detailed and precise metrics that expand on or replace the metrics we have devised.